### 2021-08-23 plos_latex_template.tex
ast reading and commenting and starting to rewrite rule1 :cookie:

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field and share our recent contributions. The larger the conference, the larger the opportunities to network and learn from your cohort. However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arend_disparity_2019, timperley_he_2020, gewin_what_2019, brown_ableism_2018}.

@@

\section{Rule X: Define what diversity and inclusion mean for your conference}

The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.

@@

As a starting point, identify which groups are usually unsupported or discriminated against in your community or event, and determine the contributing factors.

Next, set measurable indicators of success when organizing a diverse conference. The numbers resulting from these indicators are not the primary goal. Instead, this vision should guide and help hold the organizing team accountable towards these diverse and inclusive goals.

Publishing a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too. However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct, and preparing accessibility guidelines (see other Rules below).

@@

Alternatively, a conference could have a hybrid format with an in-person and online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge in this kind of setting would be to make the online component equally relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.

@@

%The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation. % rj: left this out, Heather said it doesn't make much sense here

%rj: Yani's refs to have just in case:
%https://sites.lsa.umich.edu/scottepage/wp-content/uploads/sites/344/2015/11/pnas.pdf; https://www.pnas.org/content/early/2014/11/13/1407301111; https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters

@@

Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}. People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
Inaccessible practices only exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

If the conference is in person, the venue must be accessible for people with mobility limitations. 
Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided.
Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
Regardless of the conference format --online, in person or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound. 
Platforms for conference registration and abstract submission, websites and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning for presentations should be available in more than one language if possible.
These aspects should be tested well in advance of going live.

Some practices, such as making accessible slides and presentations, will require great efforts from the presenters if they are not used to them. 
For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have.  
Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example). 
Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.

@@

To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}. The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}.%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.' 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses. A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices.

@@

When inviting people as keynote speakers, program committee, session chairs and others, we should always look for the best in their subject. Ensuring diversity in each of these roles is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups. If these groups are not showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to looking in unfamiliar places to bring in the talent.  Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}. Rather than deter us, this should thus encourage us to look further to find great people that are not routinely in the spotlight. We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.

@@

%Disabled people often repeat "Nothing about us without us". This means the actual life experiences of people from URM are not replaceable with good intentions from people outside these groups. The same holds for other dimensions of diversity. No space will be truly inclusive and welcoming to URM without people from these groups bringing their experience to bear. This means people in privileged groups may need to step down from positions of power. Of course, no-one has the same experiences and experiences the same oppressions in life. %When building a diverse team, bear in mind that

@@

Conferences, even virtual ones, should have registration fees for two primary reasons. The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee. The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events with no registration costs \cite{eventbrite_ultimate_2017}.

On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia ($\sim$ proportionate to income). A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}. Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined. You can take into account the cost of living in their countries using conversion factors (e.g. from the International Comparison Program report of the World Bank \cite{arend_disparity_2019}).

@@

%rationale: 1. they might select for the already privileged or people who dont need those to be nice on the cvs 2. they select for outgoing extroverted privileged ("brave") people 3. they are time and emotionally consuming which selects out some people 4. they single out people. even if you controlled for 1 to 3 you would "badge" people - you are creating a minority while trying to avoid it.
Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them. If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or tokenizing phrases (e.g. `grant recipient' or `diversity scholar', respectively). Do not call it a scholarship or a grant. Ultimately, it is a fee waiver or a discount for the people who need additional financial support. Also, the process for applicants should be simplified. People with limited resources already have a hard time applying for loans, grants, and scholarships. Do not make the attendees' lives more challenging by asking for long essays to convince a committee that they deserve your support. For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility. Also, the participant metadata (e.g., demographics, gender, career stage, disabilities, if they disclose them) is already a strong indicator of their potential need for aid. If the conference resources allow it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).

@@

The ten rules stated here can be adapted depending on the conference format and settings. From our own experience, we are aware that enforcing these rules requires a tremendous amount of work. We engaged in most of these practices, and others, like having an explicit payment category for postdocs or paying the whole organizing team, we learned them too late, so we share them here as part of our learning. 
Since accessibility and inclusiveness are not as widespread as it should be, the process might seem exacting. But this is where the power of communities and individuals reside. If more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead. And we will make a change towards healthier, stronger, and more inclusive communities, empowering historically marginalized people and learning from them, thereby increasing the overall quality of our conferences.%After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.

@@

% Cras egestas velit mauris, eu mollis turpis pellentesque sit amet. Interdum et malesuada fames ac ante ipsum primis in faucibus. Nam id pretium nisi. Sed ac quam id nisi malesuada congue. Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.

>>>>

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. The larger the conference, the larger the opportunities to network and learn from your cohort. However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arend_disparity_2019, timperley_he_2020, gewin_what_2019, brown_ableism_2018}.

@@

\section{Rule X: Define what diversity and inclusion mean for your conference}
%
The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.

@@

Think about the groups that are usually unsupported or discriminated against in your community or event, and about the meaning of "diverse" in your context. Imagine the result you would like to see if you succeeded in organizing a diverse conference according to your definition. Does this mean an even gender distribution in your speakers? Does this relate to people of all races -especially black people- being present as coordinators, presenters, and in the audience? Does this mean having LGBTQ+ friendly-spaces or community participation from key geographic regions? This vision should guide and help hold the organizing team accountable along the way.

Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}). %and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too. 
However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct, and preparing accessibility guidelines (see other Rules below).

%Lack of representation, unwelcoming -or overtly aggressive- environments hinder participation (or future participation) of people who could otherwise become active community members. In extreme cases, they can divert career paths, affect lives, and exclude people from some fields. So take this seriously, think about the reasons why you want to create a more diverse conference, and be open to learn. %Do not treat diversity as a checklist, do not use people from minoritized groups for image purposes (tokenization), and do not act as a savior but as a genuine collaborator. %ast: some version of this might come in a box.

@@

Alternatively, a conference could have a hybrid format with an in-person and online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge in this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.

@@

%The team may also have to review carefully and teach themselves the correct vocabulary for internal and external communication and identify the best ways to account for every culture and situation. % rj: left this out, Heather said it doesn't make much sense here %ast: we can collect part of this in rule 1, since some self-teaching, and not expecting to be taught, is necessary

%%Disabled people often repeat "Nothing about us without us", and this is true for every marginalized group. This means the actual life experiences, expertise, and insights from people in marginalized groups are not replaceable with good intentions from people outside these groups. No space will be truly inclusive and welcoming without people from these groups bringing their experience to bear. This means people in privileged groups may need to step down from positions of power. %this might
%rj: Yani's refs to have just in case:
%https://sites.lsa.umich.edu/scottepage/wp-content/uploads/sites/344/2015/11/pnas.pdf; https://www.pnas.org/content/early/2014/11/13/1407301111; https://www.mckinsey.com/business-functions/organization/our-insights/why-diversity-matters

@@

Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}. %People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
Inaccessible practices exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

If the conference is in person, the venue must be accessible for people with mobility limitations. 
Accommodations such as a quiet space for neurodivergent people and a space for breastfeeding should also be provided.
Presenters should always speak into a microphone to make it easier for the hard of hearing and for captioners or interpreters to listen to them. 
Regardless of the conference format --online, in person, or hybrid --images used in communications about the conference should be accompanied by alternative text, while videos should have both captions and sound. 
Platforms for conference registration and abstract submission, websites, and chat platforms --if used --should be screen-reader friendly and keyboard accessible, with low technology requirements (hardware, software, and internet connection). 
Captioning for presentations should be available in more than one language if possible.
These aspects should be tested well in advance of going live.

Some practices, such as making accessible slides and presentations, are not yet common practice and will require great efforts from the presenters if they are not used to them. 
For that reason, the organizing team should provide accessibility guidelines for slides and presentations, encourage their use, and be available for any questions they may have.  
Among other things, the guidelines should ask for raw and accessible material to the talks before the conference, e.g., R Markdown, HTML, or \TeX{} files; if presentations are pre-recorded, the speakers should include their video and ensure that their face is visible so that deaf and hard-of-hearing people can read their lips if needed (see \url{https://user2021.r-project.org/participation/accessibility/} for example). 
Accessibility awards at the end of the conference are an excellent way to acknowledge the speakers who were mindful of inclusiveness when preparing and delivering their talks.

@@

To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}. The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}. 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses. A diverse code of conduct team will be more understanding of intersectionality issues in discrimination and harassment practices. 
%The people who need the protection of a code of conduct are usually those with less power or privilege, as more powerful or privileged people are often already protected from most harm.'

@@

When inviting people as keynote speakers, program committee, session chairs and others, we should always look for the best in their subject. Ensuring diversity in each of these roles is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups. If these groups are not showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to looking in unfamiliar places to bring in the talent. Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}. Rather than deter us, this should encourage us to look further to find great people that are not routinely in the spotlight. We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.

@@

@@

Conferences, even virtual ones, should have registration fees for two primary reasons. The first one is that preparing the conference involves a lot of concerted effort and costs money (e.g., commercial registration tools, captioning, or conference venue, if in-person) and this translates to the participation fee. The second reason has to do with psychology -- people value more the things they pay for, and there is a lower attendance rate for free events than for events with registration costs \cite{eventbrite_ultimate_2017}.

On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}. Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined. You can take into account the cost of living in each country using conversion factors (e.g. from the International Comparison Program report of the World Bank \cite{arend_disparity_2019}). Consider that, while it is traditional for employers to provide --at least some --conference support in the Global North, this is not the case everywhere.

@@

%rationale: 1. they might select for the already privileged or people who dont need those to be nice on the cvs 2. they select for outgoing extroverted privileged ("brave") people 3. they are time and emotionally consuming which selects out some people 4. they single out people. even if you controlled for 1 to 3 you would "badge" people - you are creating a minority while trying to avoid it.
Offering scholarships or grants to attend the conference may be counterproductive, as they can be seen as more of a competition and attract people that might not need the money (e.g., when their PI could pay for it) but could apply for it to enrich their CV while filtering out people who do not feel entitled to earn them. If the goal is to help the people who really need it, the language needs to make that clear without shrouding it with fancy-sounding phrases or tokenizing phrases (e.g. `grant recipient' or `diversity scholar', respectively). Do not call it a scholarship or a grant. Ultimately, it is a fee waiver or a discount for the people who need additional financial support. Also, the process for applicants should be simplified. Applying for loans, grants, and scholarships is already an emotionally demanding task. Do not complicate the attendees' lives by asking for long essays to convince a committee that they deserve your support. For this stage of the process, a certain degree of trust goes a long way -- when they say they need the support, it is best to give them the waiver rather than second guess their eligibility. \st{Also, the participant metadata (e.g., demographics, gender, career stage, disabilities, if they disclose them) is already a strong indicator of their potential need for aid.} Other If the conference resources allow for it, you could even take further steps to offer financial support for activities that help them have the time and resources to be at the conference: child care support, transportation fees (if in-person), or internet connection services (if virtual).

@@

The ten rules stated here can be adapted depending on the conference format and settings. From our own experience, we are aware that enforcing these rules requires a tremendous amount of work. We engaged in most of these practices, and learned others too late, like having an explicit payment category for postdocs or paying the whole organizing team, so we share them here as part of our learning process. 
Since accessibility and inclusiveness are not as widespread as they should be, the process might seem exacting. But this is where the power of communities and individuals reside. If more conferences and domains employ these techniques, the process will get more streamlined, straightforward, and mainstream to adapt with minimal overhead. And we will make a change towards healthier, stronger, and more inclusive communities, empowering historically marginalized people and learning from them, thereby increasing the overall quality of our conferences.%After all, the strength of any community lies in its ability to be adaptive, inclusive, and accessible.

@@
